---
title: "Statistiek - Lab session 4"
author: "J.F.P. (Richard) (s2956586)"
date: "Generated on: `r date()`"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float: 
        collapsed: false
        smooth_scroll: true
    number_sections: true
---

# Exercise 1

```{r, fig.width=5, fig.height=5}
download.file('http://www.let.rug.nl/wieling/Statistiek-I/dat.rda','dat.rda')
load('dat.rda',verbose=T)

str(dat)
set.seed(111)
qqnorm(dat$english_score, col="blue")
qqline(dat$english_score, col="red")
boxplot(dat$english_score, main="English Score Data", col="red")
abline(h=7.5)
t.test(dat$english_score, alternative = "less", mu= 7.5 )
```

The QQ-plot shows that the distribution is normal.
The boxplot shows that most scores are lower then 7.5.
When we run the one sided t test we see that the sample mean lies higher than the actual mean.
The population is in this case a set of bachelor students who follow the course Statistiek 1.
The hypothesis was tested by a online questionairre by these same students by supplying a speech sample.

t(314) = -2.39, p =0.009

H0: $\mu$d = $\mu$t and Ha: $\mu$d < $\mu$t

We obtained English scores in a sample of 315 students of the Statistiek 1 course via an online questionnaire. 

There are 315 samples obtained through the students. Therefor there are more than 30 values and normally distributed which means the t-test could be used. 

Mean of the english: 7.347 
Significance level: 0.05
t(314) = -2.39 
p = 0.009 

The alternative hypothesis can be accepted. This means that the average english_score of the students is lower than 7.5


# Exercise 2

```{r, fig.width=5, fig.height=5}
t.test(dat$english_score, alternative = "two.sided", mu= 7.5 )
```

H0: ??d = ??t and Ha: ??d ??? ??t
This hypothesis is tested with a two-sided T-test
Significance level: = 0.05
t(314) = -2.39
p = 0.017

The results are significant because the p-level is below the significance level. Therefor H0 is rejected and we accept Ha.

H0: $\mu$d = $\mu$t
Ha: $\mu$d $\neq$ $\mu$t 


# Exercise 3

```{r, fig.width=5, fig.height=5}
table(dat$gender)
datF <- dat[dat$gender == "F", ]
datM <- dat[dat$gender == "M", ]
mean(datF$english_score)
mean(datM$english_score)
qqnorm(datF$english_score)
qqline(datF$english_score)
boxplot(datF$english_score)
qqnorm(datM$english_score)
qqline(datM$english_score)
boxplot(datM$english_score, col="red")
t.test(datF$english_score, datM$english_score)
library(effsize)
cohen.d(datF$english_score, datM$english_score)
```
We obtained English scores in a sample of 315 students of the Statistiek 1 course via an online questionnaire(222F, 93M). 

H0: $\mu$F = $\mu$M 
Ha: $\mu$F $\neq$ $\mu$M
This hypothesis is tested with a two-sided T-test
Significance level: = 0.05
p = 0.0004

The p-level is below the significance level. Therefor we must reject H0 and accept Ha.

Cohen's d: -0.4358338 (small) 
Mean english_score females: 7.203624
Mean english_score males 7.689283. 
Significance level: 0.05:

The effectsize is small, so it isn't worth mentioning. 

# Exercise 4

```{r, fig.width=5, fig.height=5}
table(dat$study)
datC <- dat[dat$study == "CIS", ]
datL <- dat[dat$study == "LING", ]
mean(datC$english_score)
mean(datL$english_score)
qqnorm(datC$english_score)
qqline(datC$english_score)
boxplot(datC$english_score, col="red")
qqnorm(datL$english_score)
qqline(datL$english_score)
boxplot(datL$english_score, col="red")
t.test(datC$english_score, datL$english_score)
cohen.d(datC$english_score,datL$english_score)
```

H0: English scores from CIS students are higher than Ling students. 
Ha: English scores from CIS students are not higher than Ling students. 

H0: $\mu$C = $\mu$L 
Ha: $\mu$C $\neq$ $\mu$L
This hypothesis is tested with a two-sided T-test
Significance level: 0.05
t(199) = 0.49
p = 0.63

The p-level is higher than the significance level. Therefor H0 will not be rejected.

Cohen's d estimate: 0.06816897 (negligible) 
Mean english_score CIS: 7.177996
Mean english_score LING 7.106976. 
Significance level: 0.05

The effectsize is negligible and there for there isn't much effect.

# Exercise 5

```{r, fig.width=5, fig.height=5}
datM <- dat[dat$gender == "M", ]
mean(datM$english_grade)
mean(datM$english_score)
qqnorm(datM$english_score-datM$english_grade)
qqline(datM$english_score-datM$english_grade)
boxplot(datM$english_score-datM$english_grade, col="red")
t.test(datM$english_score, datM$english_grade, alternative = "greater", paired = TRUE)
cohen.d(datM$english_score, datM$english_grade, paired = T)
```
H0: English scores of men are significantly equal to their English grades.
Ha: English scores of men are significantly higher than the English grades.

we are interested in assessing if for 
H0: $\mu$S = $\mu$G 
Ha: $\mu$S > $\mu$G
This hypothesis is tested with a one-sided T-test.
Significance level: 0.05:
p = 0.2478

The p-level is higher than the significance level. Therefor we do not reject H0.

Cohen's d estimate: 0.07092706 (negligible)
Mean english_grade: 7.635484
Mean LING students english_grade: 7.689283. 
Significance level: 0.05: 

The effectsize is neglible and therefor not important.

# Exercise 6

```{r, fig.width=5, fig.height=5}
datM <- dat[dat$gender == "M", ]
dat17 <- datM[datM$year == "2017", ]
mean(dat17$english_grade)
mean(dat17$english_score)
qqnorm(dat17$english_score-dat17$english_grade)
qqline(dat17$english_score-dat17$english_grade)
boxplot(dat17$english_score-dat17$english_grade, col="red")
t.test(dat17$english_score, dat17$english_grade, alternative = "greater", paired = TRUE)
cohen.d(dat17$english_score, dat17$english_grade, paired = TRUE)
```
H0: The English scores of men with the academic year of 2017 are equal than their English grades.
Ha: The English scores of men with the academic year of 2017 are significantly higher than their English grades.

H0: $\mu$S = $\mu$G 
Ha: $\mu$S > $\mu$G

This hypothesis is tested with a paired T-test.
Significance level: 0.05:
p = 0.04646

The p-level is lower than the significance level. Therefor we must reject H0 and accept Ha.

Cohen's d estimate: 0.07092706 (negligible)
Mean english_grade: 7.635484
Mean LING students english_grade: 7.689283. 
Significance level: 0.05: 

Cohen's d estimate: 0.283753 (small) 
Mean english_grade: 7.713514
Mean english_score: 7.895873. 
Significance level: 0.05: 

The effectsize is small and therefor not important.